Learning and comparing functional connectomes between populations

### 2012-12-28 paper.tex
Iter on last part

A first step to move from purely descriptive statistics to modeling with
functional connectivity analysis is to consider a correlation matrix as a
Gaussian graphical model, \emph{i.e.} a well-defined probabilistic model
that describes observed correlations in terms of an independence
structure and conditional relations \cite{lauritzen1996,varoquaux2011}.
In such settings, the inverse covariance graph or the partial
correlations are a measure of influence from one node to another, albeit
undirected.

%------------------------------------------------------------------------------
\subsection{From correlations to structural equation modeling}

\cite{mcintosh1994}
\cite{marrelec2007}
\cite{marrelec2009}

%------------------------------------------------------------------------------
\subsection{Matching model complexity to data}

@@

Setting the cursor between complex models based on a bio-physical
description, and simple phenomenological models such as correlation
matrices.

\cite{mcintosh2010}

Cross validation

>>>>

A first step to move from purely descriptive statistics to modeling with
functional connectivity analysis is to consider a correlation matrix as a
Gaussian graphical model, \emph{i.e.} a well-defined probabilistic model
that describes observed correlations in terms of an independence
structure and conditional relations \cite{lauritzen1996,varoquaux2011}.
In such settings, the inverse covariance graph or the partial
correlations are a measure of influence from one node to another, albeit
undirected. Inferring direction in a Gaussian model is impossible. Linear
structural equation models (SEMs) \cite{mcintosh1994} rely on similar
model that consists in specifying a candidate directed graphical
structure. This structure constraints the covariance matrix of the
signals and can thus be tested on observed data. In fact some forms of
SEMs are known as ``covariance structure models''. Their is thus a strong
formal link between correlation analysis in the framework of graphical
models and SEMs: the former is undirected but fully exploratory, as it
does not require the specification of candidate structure, while the
latter is directory but confirmatory. The link has been exploited to
specify candidate structures for SEMs using partial correlations
\cite{marrelec2007}. More complex models, such as dynamical causal models
(DCMs) \cite{friston2003a} or Granger causality \cite{goebel2003} require
additional hypotheses such as non-linear couplings or time lags.

Most importantly, more complex models can only be estimated to model
interactions between a smaller number of nodes. This is not only due to a
computational difficulty, but also to fundamental roadblocks in
statistics: the complexity of the model must match the richness of the
data. While injecting prior information can help model estimation, the
more informative this prior, the fragile the inference becomes. The
ongoing debate on the impact of hemodynamic lag on Granger-causality
inference \cite{smith2012a} is an example of such fragility.

@@

Setting the cursor between complex models based on a bio-physical
description, and simple phenomenological models such as correlation
matrices. Cite \cite{ravikumar2011} to argue that sparse inverse
covariance is reasonably robust to violations of the assumption of
Gaussianity.

Cross validation \cite{friston2012} to destroy the argument on the
Neyman-Pearson lemma (it is self-confirmatory).
